---
title: "SCB Analysis on Global Temps Data"
author: "Sun, Adam"
date: "February 22, 2020"
output: 
  html_document:
    theme: flatly
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
library(astsa)
library(tidyverse)
library(ggplot2)
library(scales)
library(tseries)
library(fracdiff)
library(forecast)
library(KernSmooth)
library(emmeans)
```


## SCB Prep Functions <br> 

```{r echo = FALSE}

calc_sig_hats_eq25 = function(series, p = 1/3){
  k_n = round(length(series)^p)
  breaks = c(seq(1, length(series), by = k_n))
  m = floor(length(series)/k_n)
  A_ms = cbind(idx = seq(1,length(series)), series) %>% as.data.frame() %>% 
    mutate(m = cut(idx, breaks = breaks, labels = FALSE, include.lowest = TRUE)) %>% 
    filter(!is.na(m)) %>% # we do not want remainders since not enough for one more interval
    group_by(m) %>% summarize(A_m = mean(series)) %>% select(A_m) %>% as.matrix()
  
  q3_norm = qnorm(p = 0.75, 0, 1)
  sig_hat_1 = sqrt(pi*k_n)/(2*(m-1))*sum(abs(A_ms - lag(A_ms, 1)), na.rm = TRUE)
  sig_hat_2 = sqrt(k_n/(2*q3_norm))*median(abs(A_ms - lag(A_ms, 1)), na.rm = TRUE)
  sig_hat_3 = sqrt(k_n/(2*(length(breaks)-1)))*sum((A_ms - lag(A_ms, 1))^2, na.rm = TRUE)^(1/2)
  
  return(list(k_n = k_n, A_ms = A_ms, hat_1 = sig_hat_1, hat_2 = sig_hat_2, hat_3 = sig_hat_3))
  
}


calc_k_final = function(time_idx, series, sig_hat, block_max_val = 5){
  k_ruppert = dpill(x = time_idx, y = series, blockmax = block_max_val)
  b_ruppert = k_ruppert/length(series)
  loclin_1 = locpoly(time_idx, series, bandwidth = k_ruppert, gridsize = length(time_idx))
  loclin_2 = locpoly(time_idx, series, bandwidth = k_ruppert*sqrt(2), gridsize = length(time_idx))
  e_hats = series - (2*loclin_1$y - loclin_2$y) 
  nu = sum(e_hats^2)/length(series)
  rho_hat = sig_hat^2/nu
  k_final = 2*(rho_hat)^(1/5)*k_ruppert
  return(list(k_final = k_final, e_hats = e_hats, nu = nu, rho_hat = rho_hat)) 
}


calc_suprem_rnorm = function(rand_iter, time_idx, k_final){
  loclin_iter_b_1 = locpoly(x = time_idx, y = rand_iter, bandwidth = k_final, gridsize = length(time_idx))
  loclin_iter_b_2 = locpoly(x = time_idx, y = rand_iter, bandwidth = k_final*sqrt(2), gridsize = length(time_idx))
  return(mu_suprem = max(abs(2*loclin_iter_b_1$y - loclin_iter_b_2$y))) 
}

calc_q95 = function(sup_mus_boot, CI_val = 0.95){
  return(quantile(abs(sup_mus_boot), CI_val))
} 

calc_coverage = function(time_idx, series, mean_series, q_95, sig_hat, k_final){
  interval_95 = q_95*sig_hat 
  loclin_1 = locpoly(x = time_idx, y = series, bandwidth = k_final, gridsize = length(time_idx))
  loclin_2 = locpoly(x = time_idx, y = series, bandwidth = k_final*sqrt(2), gridsize = length(time_idx))
  series_tilde = 2*loclin_1$y - loclin_2$y
  return(sum(abs(mean_series - series_tilde) > interval_95))
}

calc_mu_tilde = function(time_idx, series, k_final){
  loclin_1 = locpoly(x = time_idx, y = series, bandwidth = k_final, gridsize = length(time_idx))
  loclin_2 = locpoly(x = time_idx, y = series, bandwidth = k_final*sqrt(2), gridsize = length(time_idx))
  series_tilde = 2*loclin_1$y - loclin_2$y
  return(series_tilde)
}

calc_SCBs = function(time_idx, series, sup_mus_boot, q_95, sig_hat, k_final){
  interval_95 = q_95*sig_hat 
  series_tilde = calc_mu_tilde(time_idx = time_idx, series = series, k_final = k_final)
  temps_UB = series_tilde + interval_95
  temps_LB = series_tilde - interval_95
  return(list(q_95 = interval_95, loclin_fit = series_tilde, series_UB = temps_UB, series_LB = temps_LB))
}


```


<br>
<br>
<br>

# Data Analysis - Yearly Means <br>

## Data prep <br>

Notes: 
- We consider finding the 95% Simultaneous Convidence Bands for annual mean temps, i.e. 1660 to 2019 has 360 data points.

```{r echo = FALSE}

ce_temps_path = "https://raw.githubusercontent.com/acloudysun12/SCBs_Exploration/master/data/ce_ssn_temp_mean_raw.txt"
ce_temps = read.table(ce_temps_path, header = TRUE) # we ignore 1659 because not full year

colnames(ce_temps) = c("Year", "Qtr1", "Qtr2", "Qtr3", "Qtr4")

ce_temps_yr = rowMeans(ce_temps[,2:5], na.rm = TRUE)
df_temps_yr = cbind(ce_temps, ce_temps_yr) %>% 
  select(Year, ce_temps_yr) %>% rename(Temp = ce_temps_yr)
ce_temps_yr_ts = ts(ce_temps_yr, frequency = 1, start = 1660)
ce_temps_yr_ts = ts(ce_temps_yr, frequency = 1, start = 1660)

kn_1 = ceiling(length(ce_temps_yr_ts)^(1/3))
kn_2 = ceiling(length(ce_temps_yr_ts)^(1/2))

loclin_kn_1 = locpoly(x = time(ce_temps_yr_ts), y = ce_temps_yr_ts, kernel = "normal", 
                      gridsize = length(ce_temps_yr_ts), bandwidth = kn_1)
loclin_kn_2 = locpoly(x = time(ce_temps_yr_ts), y = ce_temps_yr_ts, kernel = "normal", 
                      gridsize = length(ce_temps_yr_ts), bandwidth = kn_2)

plot.ts(ce_temps_yr_ts, main = "Temps with bandwidth smoothers for variance estimation", ylim = c(7, 11.5))
lines(loclin_kn_1$x, loclin_kn_1$y, lty = 2, lwd = 2, col = "coral")
lines(loclin_kn_2$x, loclin_kn_2$y, lty = 2, lwd = 2, col = "blue")
legend("topleft", legend = c("data", "b = n^(1/3)", "b = n^(1/2)"), col = c(1, 2, 4), lty = c(1,2,2))

ce_resids_yr_ts_1 = loclin_kn_1$y - ce_temps_yr_ts
ce_resids_yr_ts_2 = loclin_kn_2$y - ce_temps_yr_ts

```

<br>

## Summary stats around temeprature increases

```{r}
loclin_kn_1$y[loclin_kn_1$x %in% c(1890, 1895, 1940, 1950)]

loclin_kn_2$y[loclin_kn_2$x %in% c(1890, 1895, 1940, 1950)]


```

<br>

## Estimating $\hat\sigma$ with Eq (25) <br> 

Note/Q's: <br>
- Estimate the long-term $\hat\sigma$ with the residuals of $X_i - \mu_k(i)$, where $\mu_i$ is based on local linear estimate approach with bandwidth $b \in (\frac{1}{3}, \frac{1}{2})$, as per step (a) of 4.1: Implementation. <br>
- We use the locpoly function in R which estimates local polynomial fitting with Normal kernel weights. <br>
- Applying Thm 3, we use $k_n$ = 360^(1/3) to estimate $\hat\sigma_1$, $\hat\sigma_2$, and $\hat\sigma_3$. <br>
- We round $k_n$ = 360^(1/3) to the closest whole number = `r round(length(ce_temps_yr_ts)^(1/3))`. 

```{r echo = FALSE}

sigmas = calc_sig_hats_eq25(series = ce_temps_yr_ts, p = 1/3)
sigmas$hat_3
# sigmas_test = calc_sig_hats_eq25(series = ce_temps_yr_ts - mean(ce_temps_yr_ts), p = 1/3)
# sigmas_test$hat_3

```

- The estiamted long-term SDs are `r round(sigmas$hat_1, 3)`, `r round(sigmas$hat_2, 3)`, and `r round(sigmas$hat_3, 3)` respectively. Note how $\hat\sigma_1$ and $\hat\sigma_3$ are closer to each other than to $\hat\sigma_2$. <br>
- Is the series long enough to measure the long-term variance? <br>
- We choose $\hat\sigma_3$ to adjust for our optimal bandwidth and 95% CI interval $\hat q_{95}$ because it has best asymptotic convergence as per Thm 3. However, it is somewhat more susceptible to jumps in $\mu$, which do exist in our yearly temperature series (e.g. from 1739 to 1730 where yearly temps dropped from `r ce_temps_yr_ts[time(ce_temps_yr_ts) == 1739]` to `r ce_temps_yr_ts[time(ce_temps_yr_ts) == 1739]` degrees Celsius). <br> 
- Perhaps we can test later using $\hat\sigma_2$ which is more robust to jumps in $\mu$ because it calculates $\hat\sigma$ based on the median absolute difference of the intervals. <br> 

<br>


## Calculate optimal bandwidth (Rupeprt 1995) for our mean temperature series $\mu_t$ <br>

Notes/Q's: <br>
- Use function dpill() to automatically select the bandwidth proposed by Ruppert (1995) for minimizing MSE. <br>
- Maybe also try range of bandwidths higher/lower than Ruppert ideal bandwidth for exploration. <br>
- Get variance correction factor ($\hat\rho$) for each bandwidth based on step (a) of 4.1: Implementation. <br>
- $\hat\nu$ based on the sample variance of the errors (observed - fitted local linear estimate based on bandwidth) <br>
- __Final bandwidth based on $2\hat\rho^{1/5}b_{rup}$,__ <br>
- __We also try trying $b_{rup}$ +/- 0.025 (which means length of 9).__ <br>

```{r echo = FALSE}
years = time(ce_temps_yr_ts)

k_selection = calc_k_final(years, ce_temps_yr_ts, sigmas$hat_3)
k_rup = dpill(years, ce_temps_yr_ts)
k_selection$rho_hat
k_final_0 = k_selection$k_final
k_final_lo = k_final_0 - 9
k_final_hi = k_final_0 + 9

# par(mfrow=c(1,2))
# hist(calc_k_final(years, ce_temps_yr_ts, sigmas$hat_3)$e_hats, breaks = 20, main = "errors")
# hist(rnorm(720, 0, .5), breaks = 20, main = "white noise")


# k_final_1 = calc_k_final(years, ce_temps_yr_ts, sigmas$hat_3, block_max_val = 10)$k_final
# k_final_1/k_final_0
# dpill(years, ce_temps_yr_ts, blockmax = 10)/dpill(years, ce_temps_yr_ts, blockmax = 5)

 

```
<br>

__Final "optimal" bandwidths to try based on adjustments to Ruppert optimal bandwidth:__ <br> 
__`r paste(round(c(k_final_0, k_final_lo, k_final_hi)/length(years), 3), collapse = ", ")`__ <br>
Ruppert bandwidth is `r round(k_rup/360, 3)`. $\rho$ used is `r round(k_selection$rho_hat, 3)`, $\nu$ used is `r round(k_selection$nu, 4)` <br>

<br>

## Bootstrapping to estimate the 95th percentile $\hat q_{95}$ <br>

Notes/Q's: <br>
- For each bandwidth above, we use 10^4 repetitions to bootstrap the 95% confidence interval $\hat\ q_{95}$ <br>
- From here, we get an estimate of the SCB bandwidth for each $\hat b$ as $\hat\sigma \hat{q}_{0.95}(b)$. <br> 
- Should we round bandwidth lengths to integer? Easier to interpret, but the locpoly function allows non-integer bandwidths. <br>

```{r }
set.seed(9980)

years = time(ce_temps_yr_ts)

num_sims = 10000
rnorm_sims = matrix(rnorm(num_sims*length(ce_temps_yr_ts), 0 ,1), 
                    nrow = num_sims, ncol = length(ce_temps_yr_ts))

sup_mus_boot_0 = apply(rnorm_sims, MARGIN = 1, FUN = calc_suprem_rnorm, time_idx = years, k_final = k_final_0)
sup_mus_boot_lo = apply(rnorm_sims, MARGIN = 1, FUN = calc_suprem_rnorm, time_idx = years, k_final = k_final_lo)
sup_mus_boot_hi = apply(rnorm_sims, MARGIN = 1, FUN = calc_suprem_rnorm, time_idx = years, k_final = k_final_hi)

q_95_0 = calc_q95(sup_mus_boot_0)
q_95_lo = calc_q95(sup_mus_boot_lo)
q_95_hi = calc_q95(sup_mus_boot_hi)

SCBs_k_0 = calc_SCBs(years, ce_temps_yr_ts, sup_mus_boot_0, q_95_0,sigmas$hat_3, k_final_0)
SCBs_k_lo = calc_SCBs(years, ce_temps_yr_ts, sup_mus_boot_lo, q_95_lo, sigmas$hat_3, k_final_lo)
SCBs_k_hi = calc_SCBs(years, ce_temps_yr_ts, sup_mus_boot_hi, q_95_hi, sigmas$hat_3, k_final_hi)

```

<brr>

Out of curiosity, we also look at how much of our true temperature series is covered by the SCBs. <br>
This is not grounded in theory, but perhaps it gives a sense of how strongly our temperature series fluctuates year over year. 

__Coverage of actual data with 95% SCB for bandwidth length = `r round(k_final_0)`__ -- 
`r round(1 - length(which(ce_temps_yr_ts >= SCBs_k_0$series_UB | ce_temps_yr_ts <= SCBs_k_0$series_LB))/360, 3)` <br>
__Coverage of actual data with 95% SCB for bandwidth length = `r round(k_final_lo)`__ -- 
`r round(1 - length(which(ce_temps_yr_ts >= SCBs_k_lo$series_UB | ce_temps_yr_ts <= SCBs_k_lo$series_LB))/360, 3)` <br>
__Coverage of actual data with 95% SCB for bandwidth length = `r round(k_final_hi)`__ -- 
`r round(1 - length(which(ce_temps_yr_ts >= SCBs_k_hi$series_UB | ce_temps_yr_ts <= SCBs_k_hi$series_LB))/360, 3)` <br>

<br>

## Plots with SCBs <br>

Notes/Q's: <br>
- For each bandwidth tried, we plot time series, local linear estimates, and their SCBs. 

```{r echo = FALSE}
plot(ce_temps_yr_ts, main = paste("Temps with LL est. and 95% SCB for optimal bandwidth:", round(k_final_0/360, 3)))
lines(as.numeric(years), SCBs_k_0$loclin_fit, lwd = 2, lty = 1, col = "coral")
lines(x = as.numeric(years), y = SCBs_k_0$series_UB, lwd = 2, lty = 2, col = "coral")
lines(x = as.numeric(years), y = SCBs_k_0$series_LB, lwd = 2, lty = 2, col = "coral")

plot(ce_temps_yr_ts, main = paste("Temps with LL est. and 95% SCB for bandwidth - 0.025:", round(k_final_lo/360, 3)))
lines(as.numeric(years), SCBs_k_lo$loclin_fit, lwd = 2, lty = 1, col = "coral")
lines(x = as.numeric(years), y = SCBs_k_lo$series_UB, lwd = 2, lty = 2, col = "coral")
lines(x = as.numeric(years), y = SCBs_k_lo$series_LB, lwd = 2, lty = 2, col = "coral")

plot(ce_temps_yr_ts, main = paste("Temps with LL est. and 95% SCB for bandwidth + 0.025:", round(k_final_hi/360, 3)))
lines(as.numeric(years), SCBs_k_hi$loclin_fit, lwd = 2, lty = 1, col = "coral")
lines(x = as.numeric(years), y = SCBs_k_hi$series_UB, lwd = 2, lty = 2, col = "coral")
lines(x = as.numeric(years), y = SCBs_k_hi$series_LB, lwd = 2, lty = 2, col = "coral")


```

## Fitting curves through the 95% SCBs 

With our SCBs, we can now attempt to fit different models through the temperature series. If they fall within the bounds of 95% SCB, then this suggests these trends are probable models to predict the true trend of yearly temperatures. <br> 
For the temperature data since 1850, we fit a linear model (purple) and a quadratic model (green) and observe both lines falling within the bounds of the SCBs, indicating they are possible candidates to model the temperature series. <br>
We also fit an isotonic model to the series since 1850 and note that it has 
Perhaps more interestingly, that using our 95% SCBs, we can conclude that yearly temperatures have not been constant since 1850. Since the upper limit of the 95% SCB in 1850, `r round(SCBs_k_0$series_UB[years==1850], 3)`, is smaller than the lower limit of the 95% SCB in 2019,`r round(SCBs_k_0$series_LB[years==2019], 3)`, there does not exist a horizontal line which can be drawn from 1850 to 2019 that which would be covered by the SCBs. 

```{r}

ce_temps_post_1850 = ce_temps_yr_ts[time(ce_temps_yr_ts) >= 1850] %>% ts(start = 1850)
mdl_lin = lm(ce_temps_post_1850 ~ time(ce_temps_post_1850))
summary(mdl_lin)
mdl_quad = lm(ce_temps_post_1850 ~ I(time(ce_temps_post_1850)^2) + time(ce_temps_post_1850))
summary(mdl_quad)
mdl_cubic = lm(ce_temps_post_1850 ~ 
                 I(time(ce_temps_post_1850)^3) + I(time(ce_temps_post_1850)^2) + time(ce_temps_post_1850))
summary(mdl_cubic)
0# mdl_iso = isoreg(time(ce_temps_post_1850), y = ce_temps_post_1850)

plot(ce_temps_yr_ts, main = paste("Temps with LL est, 95% SCB, and model fits since 1850"), 
     col = alpha("black", 0.5), xlim = c(1850, 2020), ylab = "Temps (Celsius)")
lines(as.numeric(years), SCBs_k_0$loclin_fit, lwd = 2, lty = 1, col = alpha("coral", 0.5))
lines(x = as.numeric(years), y = SCBs_k_0$series_UB, lwd = 2, lty = 2, col = alpha("coral", 0.5))
lines(x = as.numeric(years), y = SCBs_k_0$series_LB, lwd = 2, lty = 2, col = alpha("coral", 0.5))
lines(as.numeric(time(ce_temps_post_1850)), predict(object = mdl_lin), lwd = 2, lty = 2, col = "purple")
lines(as.numeric(time(ce_temps_post_1850)), predict(object = mdl_quad), lwd = 2, lty = 2, col = "green2")
lines(as.numeric(time(ce_temps_post_1850)), predict(object = mdl_cubic), lwd = 2, lty = 2, col = "chocolate3")
# lines(as.numeric(time(ce_temps_post_1850)), mdl_iso$yf, lwd = 2, lty = 3, col = "black")
legend("bottomright", legend = c("SCB", "Linear", "Quadratic", "Cubic"), 
       col = c("coral", "purple", "green2", "chocolate3"), lty = c(2,2,2,2))

```


# Data Analysis - Quarterly <br>

We also produce the 95% SCBs for the temperatures for each season (denoted as quarters) and see if similar conclusions about trends in the yearly temperature series can be drawn here. <br>

## Estimate $\hat\sigma$ with Eq (25) <br> 

Note/Q's: <br>
- Estimate the long-term $\hat\sigma$ with the residuals of $X_i - \mu_k(i)$ as per step (a) of 4.1: Implementation. <br>
- We use the locpoly function in R which estimates local polynomial fitting with Normal kernel weights. <br>
- Again, we use $k_n$ = 360^(1/3) to estimate $\hat\sigma_3$ = the long-term SD $\hat\sigma$. <br>

```{r echo = FALSE}

ce_temps_q1_ts = ts(ce_temps[,2], frequency = 1, start = 1660)
ce_temps_q2_ts = ts(ce_temps[,3], frequency = 1, start = 1660)
ce_temps_q3_ts = ts(ce_temps[,4], frequency = 1, start = 1660)
ce_temps_q4_ts = ts(ce_temps[,5], frequency = 1, start = 1660)

sigmas_q1 = calc_sig_hats_eq25(series = ce_temps_q1_ts, p = 1/3)
sigmas_q2 = calc_sig_hats_eq25(series = ce_temps_q2_ts, p = 1/3)
sigmas_q3 = calc_sig_hats_eq25(series = ce_temps_q3_ts, p = 1/3)
sigmas_q4 = calc_sig_hats_eq25(series = ce_temps_q4_ts, p = 1/3)

sig_hat_q1 = sigmas_q1$hat_3
sig_hat_q2 = sigmas_q2$hat_3
sig_hat_q3 = sigmas_q3$hat_3
sig_hat_q4 = sigmas_q4$hat_3

```

Results: <br>
- Not surprisingly, different quarters have varying different long-term $\hat\sigma$s. 
- The winter season (Quarter 1) has the highest SD at `r round(sig_hat_q1, 3)`. followed by spring, summer, and lastly autumn at `r round(sig_hat_q4, 3)`. <br>
- All else equal, this suggests wider bands for winter and narrower bands for autumn. <br>
- What does "blockmax" do? Why does it have such a large impact on bandwidth? <br>
- We choose a higher "blockmax" = 8 since it has such a significant effect on q3 bandwidth estimation.

<br> 

## Calculate bandwidth for temperature series by quarter <br>

```{r echo = FALSE}
years = time(ce_temps_yr_ts)

k_final_q1 = calc_k_final(years, ce_temps_q1_ts, sig_hat_q1, block_max_val = 8)$k_final
k_final_q2 = calc_k_final(years, ce_temps_q2_ts, sig_hat_q2, block_max_val = 8)$k_final
k_final_q3 = calc_k_final(years, ce_temps_q3_ts, sig_hat_q3, block_max_val = 8)$k_final
k_final_q4 = calc_k_final(years, ce_temps_q4_ts, sig_hat_q4, block_max_val = 8)$k_final

b_final_q1 = round(k_final_q1/length(years), 3)
b_final_q2 = round(k_final_q2/length(years), 3)
b_final_q3 = round(k_final_q3/length(years), 3)
b_final_q4 = round(k_final_q4/length(years), 3)

b_final_q1
b_final_q2
b_final_q3
b_final_q4

calc_sig_hats_eq25(ce_temps_q1_ts)$hat_3
calc_sig_hats_eq25(ce_temps_q2_ts)$hat_3
calc_sig_hats_eq25(ce_temps_q3_ts)$hat_3
calc_sig_hats_eq25(ce_temps_q4_ts)$hat_3

range(ce_temps_q1_ts)
range(ce_temps_q2_ts)
range(ce_temps_q3_ts) 
range(ce_temps_q4_ts)

# k_s = numeric()
# for (i in 5:15){
#   k_s = c(k_s, dpill(years, ce_temps_yr_ts, i))
# }
# q1: 26.16785 26.16785 26.16785 26.16785 26.16785 26.16785 26.16785 26.16785 26.16785 26.16785 26.16785
# q2: 8.336682 6.308946 6.308946 6.308946 4.780835 4.780835 4.780835 4.780835 4.780835 4.780835 4.780835
# q3: 37.51823 37.51823 15.66728 15.66728 12.23505 15.66728 12.23505 15.66728 15.66728 12.23505 12.23505
# q4: 11.56887 11.56887 11.56887 11.56887 11.56887 11.56887 11.56887 11.56887 11.56887 11.56887 11.56887
# yrs: 11.098656 11.098656  6.917944  6.917944  4.989034  4.989034  4.989034  4.989034  4.989034  4.989034  4.989034

```

We notice a large discrepancy in the optimal bandwidth lengths for each the temperature series by quarter. <br>
The optimal bandwidth for spring (Qtr2) is `r b_final_q1` (Ruppert's automatic bandwidth selector = `r round(dpill(years, ce_temps_q1_ts)/length(years), 3)`), while the optimal bandwidth for summer (Qtr3) is `r b_final_q3` (Ruppert bandwidth = `r round(dpill(years, ce_temps_q3_ts)/length(years), 3)`). <br>
WHY IS THIS THE CASE? Has to do with the mean of the series? If series is centered at 0, the optimal Ruppert bandwidth becomes less wide. Is this something we should always do before calculating bandwidth? Also, has to do with one setting like the "inital something". Set it to a higher number? <br>
The optimal bandwidth for winter (Qtr1) is almost 0.2, which we have shown in the simulation study to lower the nominal coverage due to higher bandwidths leading to higher bias. <br>
As such, we will be careful with analyzing the Qtr1 temperature series.

<br>

## Bootstrap SCBs and plot <br> 

We plot the quarterly temperature series along with their respective SCBs below. 

```{r echo = FALSE}

years = time(ce_temps_yr_ts)

set.seed(9975)
num_sims = 10000 
rnorm_sims = matrix(rnorm(num_sims*length(ce_temps_q1_ts), 0 ,1), 
                    nrow = num_sims, ncol = length(ce_temps_q1_ts))
sup_mus_boot_q1 = apply(rnorm_sims, MARGIN = 1, FUN = calc_suprem_rnorm, years, k_final = k_final_q1)
sup_mus_boot_q2 = apply(rnorm_sims, MARGIN = 1, FUN = calc_suprem_rnorm, years, k_final = k_final_q2)
sup_mus_boot_q3 = apply(rnorm_sims, MARGIN = 1, FUN = calc_suprem_rnorm, years, k_final = k_final_q3)
sup_mus_boot_q4 = apply(rnorm_sims, MARGIN = 1, FUN = calc_suprem_rnorm, years, k_final = k_final_q4)

SCBs_q1 = calc_SCBs(years, ce_temps_q1_ts, sup_mus_boot_q1, calc_q95(sup_mus_boot_q1), sig_hat_q1, k_final_q1)
SCBs_q2 = calc_SCBs(years, ce_temps_q2_ts, sup_mus_boot_q2, calc_q95(sup_mus_boot_q2), sig_hat_q2, k_final_q2)
SCBs_q3 = calc_SCBs(years, ce_temps_q3_ts, sup_mus_boot_q3, calc_q95(sup_mus_boot_q3), sig_hat_q3, k_final_q3)
SCBs_q4 = calc_SCBs(years, ce_temps_q4_ts, sup_mus_boot_q4, calc_q95(sup_mus_boot_q4), sig_hat_q4, k_final_q4)

```

<br>

We observe from the plots below that the local linear estiamtes and correponding 95% SCBs for the temperatures series broken out by quarter are not similar. This again is likely due to the large range of bandwidths (0.05 to 0.20) selected earlier. <br>
Like with the yearly temperature series, we fit linear and quadratic models on all four seasonal yearly temperatures since 1850. <br>
No linear nor quadratic model falls outside the range of the 95% SCB, for any quarterly temperature series. This suggests neither models are invlidated based on 95% SCB coverage alone. However, a quadratic model seems less plausible, especially for the Qtr1 temperature series since the fitted model actually show this convex curve sloping down to attain its minimum in 1897 before curving back up. <br>
A linear model seems more reasonable for all 4 quarter temperature series, and their positive slopes across all quarters further support the fact that global temperatures have been rising since 1850. 


```{r echo = FALSE}

q1_temps_post_1850 = ce_temps_q1_ts[time(ce_temps_q1_ts) >= 1850] %>% ts(start = 1850)
mdl_lin_q1 = lm(q1_temps_post_1850 ~ time(q1_temps_post_1850))
mdl_quad_q1 = lm(q1_temps_post_1850 ~ I(time(q1_temps_post_1850)^2) + time(q1_temps_post_1850))
summary(mdl_lin_q1)
summary(mdl_quad_q1)

q2_temps_post_1850 = ce_temps_q2_ts[time(ce_temps_q2_ts) >= 1850] %>% ts(start = 1850)
mdl_lin_q2 = lm(q2_temps_post_1850 ~ time(q2_temps_post_1850))
mdl_quad_q2 = lm(q2_temps_post_1850 ~ I(time(q2_temps_post_1850)^2) + time(q2_temps_post_1850))
summary(mdl_lin_q2)
summary(mdl_quad_q2)

q3_temps_post_1850 = ce_temps_q3_ts[time(ce_temps_q1_ts) >= 1850] %>% ts(start = 1850)
mdl_lin_q3 = lm(q3_temps_post_1850 ~ time(q3_temps_post_1850))
mdl_quad_q3 = lm(q3_temps_post_1850 ~ I(time(q3_temps_post_1850)^2) + time(q3_temps_post_1850))
summary(mdl_lin_q3)
summary(mdl_quad_q3)

q4_temps_post_1850 = ce_temps_q4_ts[time(ce_temps_q4_ts) >= 1850] %>% ts(start = 1850)
mdl_lin_q4 = lm(q4_temps_post_1850 ~ time(q4_temps_post_1850))
mdl_quad_q4 = lm(q4_temps_post_1850 ~ I(time(q4_temps_post_1850)^2) + time(q4_temps_post_1850))
summary(mdl_lin_q4)
summary(mdl_quad_q4)


# par(mfrow=c(1,2))

plot(ce_temps_q1_ts, xlim = c(1850, 2020), main = "Q1 temps with LL est. and SCBs",  ylab = "Temps (Celsius)")
lines(as.numeric(years), SCBs_q1$loclin_fit, lwd = 2, lty = 1, col = "coral")
lines(x = as.numeric(years), y = SCBs_q1$series_UB, lwd = 2, lty = 2, col = "coral")
lines(x = as.numeric(years), y = SCBs_q1$series_LB, lwd = 2, lty = 2, col = "coral")
lines(as.numeric(time(q1_temps_post_1850)), predict(object = mdl_lin_q1), lwd = 2, lty = 2, col = "purple")
lines(as.numeric(time(q1_temps_post_1850)), predict(object = mdl_quad_q1), lwd = 2, lty = 2, col = "green2")
legend("bottomright", legend = c("SCB", "Linear", "Quadratic"), 
       col = c("coral", "purple", "green2"), lty = c(2,2,2), cex = 0.75)

plot(ce_temps_q2_ts, xlim = c(1850, 2020), main = "Q2 temps with LL est. and SCBs", ylab = "Temps (Celsius)")
lines(as.numeric(years), SCBs_q2$loclin_fit, lwd = 2, lty = 1, col = "coral")
lines(x = as.numeric(years), y = SCBs_q2$series_UB, lwd = 2, lty = 2, col = "coral")
lines(x = as.numeric(years), y = SCBs_q2$series_LB, lwd = 2, lty = 2, col = "coral")
lines(as.numeric(time(q2_temps_post_1850)), predict(object = mdl_lin_q2), lwd = 2, lty = 2, col = "purple")
lines(as.numeric(time(q2_temps_post_1850)), predict(object = mdl_quad_q2), lwd = 2, lty = 2, col = "green2")
legend("bottomright", legend = c("SCB", "Linear", "Quadratic"), 
       col = c("coral", "purple", "green2"), lty = c(2,2,2), cex = 0.75)

plot(ce_temps_q3_ts, xlim = c(1850, 2020), main = "Q3 temps with LL est. and SCBs", ylab = "Temps (Celsius)")
lines(as.numeric(years), SCBs_q3$loclin_fit, lwd = 2, lty = 1, col = "coral")
lines(x = as.numeric(years), y = SCBs_q3$series_UB, lwd = 2, lty = 2, col = "coral")
lines(x = as.numeric(years), y = SCBs_q3$series_LB, lwd = 2, lty = 2, col = "coral")
lines(as.numeric(time(q2_temps_post_1850)), predict(object = mdl_lin_q3), lwd = 2, lty = 2, col = "purple")
lines(as.numeric(time(q2_temps_post_1850)), predict(object = mdl_quad_q3), lwd = 2, lty = 2, col = "green2")
legend("bottomright", legend = c("SCB", "Linear", "Quadratic"), 
       col = c("coral", "purple", "green2"), lty = c(2,2,2), cex = 0.75)

plot(ce_temps_q4_ts, xlim = c(1850, 2020), main = "Q4 temps with LL est. and SCBs", ylab = "Temps (Celsius)")
lines(as.numeric(years), SCBs_q4$loclin_fit, lwd = 2, lty = 1, col = "coral")
lines(x = as.numeric(years), y = SCBs_q4$series_UB, lwd = 2, lty = 2, col = "coral")
lines(x = as.numeric(years), y = SCBs_q4$series_LB, lwd = 2, lty = 2, col = "coral")
lines(as.numeric(time(q2_temps_post_1850)), predict(object = mdl_lin_q4), lwd = 2, lty = 2, col = "purple")
lines(as.numeric(time(q2_temps_post_1850)), predict(object = mdl_quad_q4), lwd = 2, lty = 2, col = "green2")
legend("bottomright", legend = c("SCB", "Linear", "Quadratic"), 
       col = c("coral", "purple", "green2"), lty = c(2,2,2), cex = 0.75)


```

<br>

Test whether the coefficients of linear trends are statistically significantly different. ANOVA

```{r}

ce_temps_post_1850_long = ce_temps %>% gather(key = Qtr, value = Temp, -Year) %>% filter(Year >= 1850)

mdl_temps_post_1850 = lm(Temp ~ Year*Qtr, data = ce_temps_post_1850_long)

summary(mdl_temps_post_1850)

anova(mdl_temps_post_1850)

slopes = emtrends(mdl_temps_post_1850, 'Qtr', var = 'Year') #gets each slope
slopes

pairs(slopes)


```

<br>

Is flat temperatures possible for annual temperatures of each season? i.e. Can temperatures be not changing under our 95% SCBs?

```{r echo = TRUE}

print("Yearly avg annual temperatures")
min(round(SCBs_k_0$series_UB[years>=1850], 3))
max(round(SCBs_k_0$series_LB[years>=1850], 3))
min(round(SCBs_k_0$series_UB[years>=1850], 3)) > max(round(SCBs_k_0$series_LB[years>=1850], 3))
min(round(SCBs_k_0$series_UB[years>=1850], 3)) - max(round(SCBs_k_0$series_LB[years>=1850], 3))

print("Q1 annual temperatures")
min(round(SCBs_q1$series_UB[years>=1850], 3))
max(round(SCBs_q1$series_LB[years>=1850], 3))
min(round(SCBs_q1$series_UB[years>=1850], 3)) > max(round(SCBs_q1$series_LB[years>=1850], 3))
min(round(SCBs_q1$series_UB[years>=1850], 3)) - max(round(SCBs_q1$series_LB[years>=1850], 3))

print("Q2 annual temperatures")
min(round(SCBs_q2$series_UB[years>=1850], 3))
max(round(SCBs_q2$series_LB[years>=1850], 3))
min(round(SCBs_q2$series_UB[years>=1850], 3)) > max(round(SCBs_q2$series_LB[years>=1850], 3))
min(round(SCBs_q2$series_UB[years>=1850], 3)) - max(round(SCBs_q2$series_LB[years>=1850], 3))

print("Q3 annual temperatures")
min(round(SCBs_q3$series_UB[years>=1850], 3))
max(round(SCBs_q3$series_LB[years>=1850], 3))
min(round(SCBs_q3$series_UB[years>=1850], 3)) > max(round(SCBs_q3$series_LB[years>=1850], 3))
min(round(SCBs_q3$series_UB[years>=1850], 3)) - max(round(SCBs_q3$series_LB[years>=1850], 3))

print("Q4 annual temperatures")
min(round(SCBs_q4$series_UB[years>=1850], 3))
max(round(SCBs_q4$series_LB[years>=1850], 3))
min(round(SCBs_q4$series_UB[years>=1850], 3)) > max(round(SCBs_q4$series_LB[years>=1850], 3))
min(round(SCBs_q4$series_UB[years>=1850], 3)) - max(round(SCBs_q4$series_LB[years>=1850], 3))

```


<br> 

Playground/Additional Notes <br>

- Thoughts with fake data. Test structural breaks etc, like suppose we added 2 degrees Celsius to all temps after 1990, etc.

```{r echo = FALSE}
ce_temps_yr_ts_fake = ifelse(time(ce_temps_yr_ts) >= 1900, ce_temps_yr_ts + 2, ce_temps_yr_ts)
plot(ce_temps_yr_ts_fake)
lines(lowess(ce_temps_yr_ts_fake, f= 0.1), lwd = 2, col = 2)
lines(lowess(ce_temps_yr_ts_fake), lwd = 2, col = 4)

```

<br> 

<!-- ## Calculate coverage of actual data with 95% SCB for temperature series<br>  -->

```{r eval = FALSE, echo = FALSE}
print("Empirical coverage for Q1 temps: ")
round(1 - length(which(ce_temps_q1_ts >= SCBs_q1$series_UB | 
                         ce_temps_q1_ts <= SCBs_q1$series_LB))/360, 3)

print("Empirical coverage for Q2 temps: ")
round(1 - length(which(ce_temps_q2_ts >= SCBs_q2$series_UB | 
                         ce_temps_q2_ts <= SCBs_q2$series_LB))/360, 3)

print("Empirical coverage for Q3 temps: ")
round(1 - length(which(ce_temps_q3_ts >= SCBs_q3$series_UB | 
                         ce_temps_q3_ts <= SCBs_q3$series_LB))/360, 3)

print("Empirical coverage for Q4 temps: ")
round(1 - length(which(ce_temps_q4_ts >= SCBs_q4$series_UB | 
                         ce_temps_q4_ts <= SCBs_q4$series_LB))/360, 3)

```

<br>
<br>


<!-- ## Estimating $\hat\sigma^2$ with SDF <br> -->

<!-- Notes/Q's: <br> -->
<!-- - We could use spectral density function to get estimate for $\gamma(0)$ aka $\hat\sigma^2$. <br> -->
<!-- - What smoothing should be used for periodogram? <br> -->
<!-- - what type of smoothing for periodogram? Modified Daniell okay? Looked at no smoothing, modified.daniell.c(4,4) (0.03 bandwidth), and modified.daniell.c(18,18) (0.15 banwidth). Seems sigma_sq_hat estimate is similar across any smoothing. We pick the largest estimate. <br> -->
<!-- - Estimate of sigma_sq_hat should be 2*(area of periodogram), since it's calculated in range of $w \in (0, 1/2)$ but we want range  $w \in (-1/2, 1/2)$? <br> -->
<!-- - Is my periodogram approach correct? Why is there mention of the spectral density function = $\sigma^2/(2\pi)$ in Part 5. oe this mean I shouldd multiply my periodogram estimate by $2\pi$? <br> -->
<!-- - Should we use periodogram appraoch for variance estimate? -->

```{r eval = FALSE, echo = FALSE}

# Note ce_resids_yr_ts_1 is not used for estimating sigma_sq_hat since ce_resids_yr_ts_2 bigger

smooth_1 = kernel("modified.daniell", c(4,4))
round(1/(sum(smooth_1$coef^2) + sum(smooth_1$coef[-1]^2))/length(ce_temps_yr_ts), 3)

smooth_2 = kernel("modified.daniell", c(18,18))
round(1/(sum(smooth_2$coef^2) + sum(smooth_2$coef[-1]^2))/length(ce_temps_yr_ts), 3)

Iw_smooth_0 = mvspec(ce_resids_yr_ts_2, log = "no") # this uses no smoothing of periodogram
Iw_smooth_1 = mvspec(ce_resids_yr_ts_2, kernel = smooth_1, log = "no") # this uses modified Daniell
Iw_smooth_2 = mvspec(ce_resids_yr_ts_2, kernel = smooth_2, log = "no") # this uses modified Daniell
sig_sq_hat_0 = 2*sum(Iw_smooth_0$freq[1] * Iw_smooth_0$spec)
sig_sq_hat_0
sig_sq_hat_1 = 2*sum(Iw_smooth_1$freq[1] * Iw_smooth_1$spec)
sig_sq_hat_1
sig_sq_hat_2 = 2*sum(Iw_smooth_2$freq[1] * Iw_smooth_2$spec)
sig_sq_hat_2

```

